Quantitative Benchmarks for Theory Exploration

### 2017-05-05 PhDSymposium2017/abstract/abstract.tex
Final fixes

@@

\begin{document}
\maketitle
\begin{abstract}
  \emph{Automated theory exploration} has the potential to help scientists,
  mathematicians and programmers discover unexpected properties of their models,
  theories and programs; but has so far been limited to small-scale experiments,
  with only a few, hand-picked, data points. We address this problem by
  re-purposing existing benchmarks from the theorem proving domain, producing an
  order of magnitude more data for evaluation, and use it to evaluate the
  performance of a state of the art exploration system. To combat the long run
  times of these systems, we also propose a pre-processing method to identify
  promising subsets of a theory, sacrificing completeness to gain speed. It is
  our hope that the availability of robust evaluation methods, and scalable
  algorithms, will encourage competition and innovation, making theory
  exploration more useful and relevant as a formal methods tool.
\end{abstract}

@@

@@

@@

@@

Existing ATE programs have been evaluated this way on theories of lists and
natural numbers taken from the Isabelle proof
assistant~\cite{claessen2013automating}. The discovered properties are compared
against those in Isabelle, to see how many match up. Unfortunately these are the
only data sets used, and each is rather small (containing 9 and 12 properties,
respectively). This is not enough data from which to draw meaningful comparisons
or extrapolations.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{runtimes}
  \caption{Median runtime of the QuickSpec ATE system for 30 iterations on
    theories of various size, sampled from TIP. Error bars show the
    interquartile range. We time out at 1000 seconds.}\label{fig:runtimes}
\end{figure}

Figure~\ref{fig:runtimes} shows the median run time of the existing
\textsc{QuickSpec}\cite{QuickSpec} system as we explore theories of various size
sampled from our benchmark. We have found that the time taken can vary greatly,
even among theories of the same size, with most finishing after a few seconds
but some taking hours; this can be seen by the increase in range for larger
theories, despite the median remaining low.

This variance makes current ATE systems inconvenient for widespread use. Finding
a pattern in these long-running inputs is current work; our hope is to make such
tools more predictable and tractable using approaches like our relevance filter.

>>>>

\setlength{\belowcaptionskip}{-10pt}

@@

\begin{document}
\maketitle
\begin{abstract}
  \emph{Automated theory exploration} has the potential to help scientists,
  mathematicians and programmers discover unexpected properties of their models,
  theories and programs; but has so far been limited to small-scale experiments,
  with only a few, hand-picked, data points. We address this problem by
  re-purposing existing benchmarks from the theorem proving domain, producing an
  order of magnitude more data for evaluation, and use it to evaluate the
  performance of a state of the art exploration system. To combat the long run
  times of these systems, we also propose a pre-processing method to identify
  promising subsets of a theory, sacrificing completeness to gain speed. It is
  our hope that the availability of robust evaluation methods, and scalable
  algorithms, will encourage competition and innovation, making theory
  exploration more useful and relevant in formal methods.
\end{abstract}

%\vspace{-2mm}

\textbf{Keywords:} Theory exploration, automated reasoning, automated theorem
proving

%\vspace{-2mm}

@@

\vspace{-2mm}

@@

\vspace{-5mm}

@@

\vspace{-3mm}

@@

Existing ATE programs such as \textsc{QuickSpec}~\cite{QuickSpec} and
\textsc{IsaCoSy}~\cite{johansson2009isacosy} have been evaluated this way on
theories of lists and natural numbers taken from the Isabelle proof
assistant~\cite{claessen2013automating}. The discovered properties are compared
against those in Isabelle, to see how many match up. Unfortunately these are the
only data sets used, and each is rather small (containing 9 and 12 properties,
respectively). This is not enough data from which to draw meaningful comparisons
or extrapolations.

@@

\begin{figure}
  \centering
  \includegraphics[width=0.45\textwidth]{runtimes}
  \vspace{-3mm}
  \caption{Median runtime of the \textsc{QuickSpec} ATE system for 30 iterations
    on theories of various size, sampled from TIP. Error bars show the
    interquartile range. We time out at 1000 seconds.}\label{fig:runtimes}
  \vspace{-5mm}
\end{figure}

Figure~\ref{fig:runtimes} shows the median run time of \textsc{QuickSpec} as we
explore theories of various size sampled from our benchmark. We have found that
the time taken can vary greatly, even among theories of the same size, with most
finishing after a few seconds but some taking hours; this can be seen by the
increase in range for larger theories, despite the median remaining low.

This variance makes current ATE systems inconvenient for widespread use. Finding
a pattern in these long-running inputs is current work; our hope is that
approaches like our relevance filter make ATE more predictable and tractable.
